Parameters estimation via HOSVD

### 2024-10-12 Code/R/Research/HOSVD parameters estimate.Rmd
Added 1d methods comparison

```{r tensor-based-esprit}
tens_esprit <- function(s, I, L, groups, kind = c("HO-SSA", "HO-MSSA"), est_dim = 3, status = TRUE) {
  if (identical(kind[1], "HO-SSA"))
    H <- tens3(s, I, L)
  else if (identical(kind[1], "HO-MSSA"))
  {
    if (missing(L))
      L <- I
    H <- tens3(s, L, kind = "HO-MSSA")
  }
  max_rank <- max(sapply(groups, max))
  H.hooi <- tucker_mod(H, rep(max_rank, 3), status = status)
  estimates <- list()
  for (i in seq(groups)) {
    U <- H$U[[est_dim]][, groups[[i]], drop = FALSE]
    Z <- solve(U[-nrow(U), ], U[-1, ])
    poles <- eigen(Z, only.values = TRUE)$values
    estimates[[i]] <- list(
      poles = poles,
      damping = -Re(log(poles)),
      freq = Im(log(poles)) / 2 / pi
    )
  }
  estimates
}
```

```{r test-1d-esprit-no-noise}

```

@@

```{r test-1d-esprit-with-noise}
set.seed(5)
n <- 25
SNR <- 30
signal <- exp((-0.01 + 2i * pi * 0.2) * 0:(n - 1)) + exp((-0.02 + 2i * pi * 0.22) * 0:(n - 1))
sd <- sqrt(mean(abs(signal)^2) / SNR)
noise <- CCSWGN(n, mean = 0, sd = sd)
x <- signal + noise
```

>>>>

```{r tensor-based-esprit}
tens_esprit <- function(s, I, L, groups, kind = c("HO-SSA", "HO-MSSA"), est_dim = 3, status = TRUE) {
  if (identical(kind[1], "HO-SSA"))
    H <- tens3(s, I, L)
  else if (identical(kind[1], "HO-MSSA"))
  {
    if (missing(L))
      L <- I
    H <- tens3(s, L, kind = "HO-MSSA")
  }
  max_rank <- max(sapply(groups, max))
  H.hooi <- tucker_mod(H, rep(max_rank, 3), status = status)
  estimates <- list()
  for (i in seq(groups)) {
    U <- H.hooi$U[[est_dim]][, groups[[i]], drop = FALSE]
    Z <- qr.solve(U[-nrow(U), ], U[-1, ])
    poles <- eigen(Z, only.values = TRUE)$values
    estimates[[i]] <- list(
      poles = poles,
      rates = Re(log(poles)),
      frequencies = Im(log(poles)) / 2 / pi
    )
  }
  estimates
}
```

```{r test-1d-ssa-esprit-no-noise}
n <- 25
x <- exp((-0.01 + 2i * pi * 0.2) * 0:(n - 1)) + exp((-0.02 + 2i * pi * 0.22) * 0:(n - 1))
L <- 15
x.ssa <- ssa(x, L = L) 
pars <- parestimate(x.ssa, groups = list(1:2))
cat("Rates:\n", pars$rates, "\nFrequencies:\n", pars$frequencies)
```
```{r test-1d-tens-esprit-no-noise}
n <- 25
x <- exp((-0.01 + 2i * pi * 0.2) * 0:(n - 1)) + exp((-0.02 + 2i * pi * 0.22) * 0:(n - 1))
I <- 14; L <- 8
pars <- tens_esprit(x, I, L, groups = list(1:2), status = FALSE)
cat("Rates:\n", pars[[1]]$rates, "\nFrequencies:\n", pars[[1]]$frequencies)
```

@@

```{r rrmse}
rrmse <- function(true, estimate) {
  100 * sqrt(mean(abs(true - estimate)^2)) / true
}
```

```{r test-1d-esprit-noise}
set.seed(10)
n <- 25
L <- 15
I <- 14; L.tens <- 8
rates.true <- c(-0.01, -0.02)
freqs.true <- c(0.2, 0.22)
signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) +
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))
sd <- 0.025
R <- 1000

pars.ssa <- list(
  rate1 = numeric(R),
  rate2 = numeric(R),
  freq1 = numeric(R),
  freq2 = numeric(R)
)
pars.tssa <- list(
  rate1 = numeric(R),
  rate2 = numeric(R),
  freq1 = numeric(R),
  freq2 = numeric(R)
)
pb <- txtProgressBar(max = R, style = 3)
for (i in seq(R)) {
  noise <- CCSWGN(n, mean = 0, sd = sd)
  x <- signal + noise
  x.ssa <- ssa(x, L = L)
  par.ssa <- parestimate(x.ssa, groups = list(1:2))
  pars.ssa$rate1[i] <- max(par.ssa$rates[1], par.ssa$rates[2])
  pars.ssa$rate2[i] <- min(par.ssa$rates[1], par.ssa$rates[2])
  pars.ssa$freq1[i] <- min(par.ssa$frequencies[1], par.ssa$frequencies[2])
  pars.ssa$freq2[i] <- max(par.ssa$frequencies[1], par.ssa$frequencies[2])
  par.tssa <- tens_esprit(x, I, L.tens, groups = list(1:2), status = FALSE)
  pars.tssa$rate1[i] <- max(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
  pars.tssa$rate2[i] <- min(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
  pars.tssa$freq1[i] <- min(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])
  pars.tssa$freq2[i] <- max(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])
  setTxtProgressBar(pb, i)
}
close(pb)
cat("Rate 1, SSA: ", rrmse(rates.true[1], pars.ssa$rate1), '\n')
cat("Rate 1, HO-SSA: ", rrmse(rates.true[1], pars.tssa$rate1), '\n')
cat("Rate 2, SSA: ", rrmse(rates.true[2], pars.ssa$rate2), '\n')
cat("Rate 2, HO-SSA: ", rrmse(rates.true[2], pars.tssa$rate2), '\n')
cat("Freq 1, SSA: ", rrmse(freqs.true[1], pars.ssa$freq1), '\n')
cat("Freq 1, HO-SSA: ", rrmse(freqs.true[1], pars.tssa$freq1), '\n')
cat("Freq 2, SSA: ", rrmse(freqs.true[2], pars.ssa$freq2), '\n')
cat("Freq 2, HO-SSA: ", rrmse(freqs.true[2], pars.tssa$freq2), '\n')
```
